---
title: "Using the CreateBzoneDev Module"
author: "Brian Gregor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the CreateBzoneDev Module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **CreateBzoneDev** module creates several Bzone development characteristics including the *PopDen* (population density in persons per square mile), *DistFromCtr* the distance in miles from the urban area center (for metropolitan Bzones only), *SfdNum* the number of households in single-family dwellings, and *MfdNum* the number of households in multifamily dwellings. The model is run in the framework with the following command:
```{r, eval=FALSE}
runModule(ModuleName = "CreateBzoneDev", PackageName = "vedemo1", 
          Year = XXXX, IgnoreInp_ = NULL, IgnoreSet_ = NULL)
```
where the a four-digit year value is substituted for `XXXX`. Most often the Year argument will be input programmatically. The IgnoreInp_ and IgnoreSet_ arguments will ignore any specified 'input' and/or 'set' specifications identified by the specification names. This option is provided to allow knowledgeable users to substitute for specified inputs and/or outputs. It should only be used with extreme caution.

The *CreateBzoneDev* module must be run after the *CreateBzones* module has been run. The module is run by Marea (the RunBy component is assigned the value "Marea"). This is done because the distribution of Bzone population densities is dependent on the overall Marea area (square miles).

The *CreateBzoneDev* module depends on one scenario input that is provided at the Marea level. The input file must be named "marea_area.csv" and must be located in the "inputs" directory of the model run setup. An example is included in the "tests/data" directory of this package. The structure of the file is as follows:
```{r, echo=FALSE, results='asis'}
MareaArea_df <- read.csv("../tests/data/marea_area.csv", as.is = TRUE)
pander::pandoc.table(MareaArea_df)
```
The column names for the file must be as shown. Rows need to exist for all Mareas and all years for which the model will be run. The values in the *Area* column represent the area in square miles of the urbanized area. Values may be integer or and decimal values. A period must be used to represent the decimal point. Thousands separators may not be used.

This module uses three model parameter files that are located in the "inst/extdata" directory of this package: *pop_by_distance.csv*, *pop_density_by_distance.csv*, and *misc_dev_parameters.csv*. The accompanying *pop_by_distance.txt*, *pop_density_by_distance.txt*, and *misc_dev_parameters.txt* files describes the source and structure of those files. A substitute for any or all of the files (with the same names and structures) may be used instead to modify the module so that it is more representative of the region and/or base year for the model deployment. If you wish to do this, you must obtain the source package. Then modify the files, making certain that the modifications are consistent with the specifications described in the corresponding description files. Then build the package to create a binary version. The best way to do this is using the RStudio IDE for R development along with the *devtools* package. RStudio is free and it along with *devtools* has everything you need to easily create a binary package from a source package. The `devtools::build(binary = TRUE)` will take care of things. For more information about R packages, I recommend referring to Hadley Wickham's book, *R Packages*, and/or to his [website](http://r-pkgs.had.co.nz/).
