---
title: "Documentation for Demo Module"
author: "Brian Gregor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Documentation for Household Synthesis Module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This file documents the demo module for doing nothing. 
```{r}
# Load functions and data used in this vignette
source("synthesize_hh/functions.r")
PrsnInput_df <- read.csv("synthesize_hh/pums_person.csv", as.is=TRUE)
HhInput_df <- read.csv("synthesize_hh/pums_housing.csv", as.is=TRUE)
```

The number of household types can be reduced substantially by removing a small portion of the households. This can be seen by comparing the number of household types when all households are selected vs. when the default of 99% of households are selected.
```{r}
HtProb1_HtAp <- calcAgeProbByHhType(PrsnInput_df, HhInput_df, Threshold=1)
HtProb2_HtAp <- calcAgeProbByHhType(PrsnInput_df, HhInput_df, Threshold=0.99)
```
When all households are selected, there are `r nrow(HtProb1_HtAp)` household types. Setting the selection threshold to 0.99 reduces the number of household types to `r nrow(HtProb2_HtAp)`.

The `calcAgeProbByHhType` function tabulates the proportions of persons in each age group by household type. It produces a matrix where the rows correspond to the selected household types and the columns correspond to the 6 age groups. The values in each column add up to one.

The `createHhByAge` function uses the matrix created by `calcAgeProbByHhType` and a demographic forecast of the number of persons in each age group to produce a set of "synthetic households" that is consistent with the demographic forecast and reflects the household type distribution. The matrix created by `calcAgeProbByHhType` is used to establish the initial allocation of population to household types. However, since the household type probabilities by age group are calculated as though household type is a characteristic of individual persons rather than a joint characteristic of multiple persons in the household, the assignment of persons to households will not be internally consistent. For example, suppose that 1000 persons ages 0-14 and 1200 persons aged 20-29 are assigned to a household type in which there are two persons in each of those age groups in the household. In this case there would be two different estimates of the number of households of this type (500 vs. 600 households). The function uses an iterative proportional fitting (IPF) procedure to reconcile such inconsistencies so that a consistent set of households is produced that matches the population forecast. The steps in procedure started at the beginning are as follows:  

1. The number of persons by household type and age group is divided by a matrix of specified number of persons by age group for each household type. This produces a matrix of the number of households by household type and age group.

2. The mean number of households is calculated for each household type (excluding zeros) to reconcile conflicting estimates of households by household type.

3. The mean number of households by household type is multiplied by the specified number of persons by age group for each household type to produce revised estimates of the number of persons by age group and household type. The age group proportions are updated using this matrix of values.

4. The total of persons by age group is calculated from the revised estimates of the number of persons by age group and household type. These age group totals are compared to the input forecast of population by age group. If all of the totals are no more than 0.1% different, the procedure is terminated. If any are greater, the procedure continues by calculating the difference between the input forecast of population by age group and the age group totals of the revised estimates.

5. The differences in population by age group are allocated to household types using the updated age group proportions calculated in the 3rd step. The resulting matrix is added to the revised estimate households by age group and housing type to produce an updated estimate and the procedure starts over.

After completing the balancing process, the resulting estimates of households by household type are rounded to whole numbers. This then becomes the basis for creating the synthetic households with each household type being replicated by the number of households assigned to it. In addition to being assigned a number of people by age group, each household is assigned a unique identification number and a value for the household size. The order of households is randomized. Following is an example of what the synthetic households look like.

```{r}
Prsn2000_CoAp <- as.matrix(read.csv("synthesize_hh/pop_by_age_2000.csv", as.is=TRUE, row.names=1))
Prsn2000_Ap <- colSums(Prsn2000_CoAp)
SynPop_df <- data.frame(createHhByAge(Prsn2000_Ap, HtProb2_HtAp))
head(SynPop_df)
```


