---
title: "Using the CreateBzones Module"
author: "Brian Gregor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the CreateBzones Module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **CreateBzones** module creates a set of simulated Bzones which represent typical census block groups and reflect scenario inputs regarding the proportion allocation of Azone households to three development types: Metropolitan, Town, and Rural. A number of households is allocated to each Bzone. The model is run in the framework with the following command:
```{r, eval=FALSE}
runModule(ModuleName = "CreateBzones", PackageName = "vedemo1", 
          Year = XXXX, IgnoreInp_ = NULL, IgnoreSet_ = NULL)
```
where the a four-digit year value is substituted for `XXXX`. Most often the Year argument will be input programmatically. The IgnoreInp_ and IgnoreSet_ arguments will ignore any specified 'input' and/or 'set' specifications identified by the specification names. This option is provided to allow knowledgeable users to substitute for specified inputs and/or outputs. It should only be used with extreme caution.

The *CreateBzones* module must be run after the *CreateHouseholds* module and before the *CreateBzoneDev* modules are run. The module is run for the entire model area at once (the RunBy component is assigned the value "All). This is done because the full set of Bzones must be created in order to determine how many rows the *Bzone* table will have. This needs to be determined in order to initialize the table.

The *CreateBzones* module depends on one scenario input file that provides values for the proportions of households in metropolitan, town and rural development types at the Azone level. The input file must be named "devtype_proportions.csv" and must be located in the "inputs" directory of the model run setup. An example is included in the "inst/extdata" directory of this package. The structure of the file is as follows:
```{r, echo=FALSE, results='asis'}
DevTypeProp_df <- read.csv("../inst/extdata/devtype_proportions.csv", as.is = TRUE)
pander::pandoc.table(DevTypeProp_df)
```
The column names for the file must be as shown. Rows need to exist for all Azones and all years for which the model will be run. The data in the *Metropolitan*, *Town*, and *Rural* columns must be decimal values. A period must be used to represent the decimal point. The three values must add to 1.

This module uses no parameter files so there is no need for customizing this module to reflect conditions in the region where the model is being applied.
